---
title: "hospital"
author: "xj2249"
date: "2019/10/31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(faraway)
library(broom)
```

```{r}
data_hosp <- 
        read.csv("./Hospital.csv") %>% 
        janitor::clean_names() %>% 
        mutate(region = factor(region),
               ms = ifelse(medschl == 1,1,ifelse(medschl == 2, 0, NA))
               )
names(data_hosp)
```

# Simple linear regression: Length of stay (LOS) vs number of BEDS 
```{r}
reg_hos<-lm(data_hosp$los~data_hosp$beds)
summary(reg_hos)
```

## Matrix model
```{r}
model.matrix(reg_hos)
model.matrix(reg_hos) %>% head
```

# Multiple linear regression
```{r}
# Var 1: Number of BEDS 
# Var 2: INFRISK (prob. % of getting an infection during hospitalization)

regmult1_hos<-lm(data_hosp$los~data_hosp$beds + data_hosp$infrisk)

# Analyze the regression results
summary(regmult1_hos)
```


## Multiple linear regression: BEDS and INFRISK and NURSE
```{r}
regmult2_hos<-lm(los ~ beds + infrisk + nurse,data = data_hosp)
summary(regmult2_hos)
```

## Categorical predictor REGION: multiple levels
```{r}
# Simple linear regression with predictor REGION (1:NE, 2:NC, 3:S, 4:W)
data_hosp %>% lm(los~region, data = .) %>% summary

# Change the reference category for REGION (from 1 to 3)
# Intercept added
data_hosp %>% 
        mutate(region = relevel(region,ref = 3)) %>% 
        lm(los~region, data = .) %>% 
        summary

# No intercept model
data_hosp %>% 
        lm(LOS~0+region, data = .) %>% 
        summary

```


## Multiple linear regression: INFRISK, new MS and Region (1:NE, 2:NC, 3:S, 4:W)
```{r}
regmult4_hos <- 
        data_hosp %>% 
        lm(los~infrisk + ms+ region, data = .)
summary(regmult4_hos)
```

### 'General' global test for all predictors
```{r}
anova(regmult4_hos)
```

### try jeff's way to test region
```{r}
reg4_null <- 
        data_hosp %>% 
        lm(los~infrisk + ms, data = .)
anova(reg4_null,regmult4_hos)
# results for "region" variable is the same. we're both test if the beta for this variable is zero. (?)
```


## Multiple linear regression: new MS, Region (1:NE, 2:NC, 3:S, 4:W) and their interaction
```{r}
regmult5_hos <- lm(los~infrisk*ms, data = data_hosp)
summary(regmult5_hos)
```

### Vizualize interaction for reg5
```{r}
gen_risk <- seq(0,8,0.1)
beta <- regmult5_hos$coefficients
yhat0 <- beta[1] + beta[2]*gen_risk
yhat1 <- beta[1] + beta[3] + (beta[2] + beta[4])*gen_risk

plot(data_hosp$infrisk, data_hosp$los, 
     xlab = 'Infection Risk (%)', ylab = 'Length of stay (days)',
     main = ('Scatter plot with overlaid fitted lines by medical school affiliation'))
lines(gen_risk, yhat0, lwd = 2, col = 2)
lines(gen_risk, yhat1, lwd = 2, col = 3)
```




