---
title: "P8130_hw2"
author: "xj2249"
date: "2019/10/1"
output: 
  pdf_document:
  latex_engine: xelatex
---

# Problem1
## a)
Let X denote the number of people who develop uveal melanoma in a given year.
Since n is very large and p is small, X follows a poisson distribution as followed:
$$
X\thicksim Poi(42.5),  \quad \lambda = 42.5
$$

Therefore
$$
P(X=30) = \frac{\lambda^xe^{- \lambda}}{x!} =  \frac{42.5^{30}e^{-42.5}}{30!} = `r dpois(30,42.5)` \approx0.009
$$

## b)
The population of Asians, non-Hispanic Whites and Black are `r  8500000*c(0.14,0.428,24.3)`, 
and therefore, $X_{Asians}$,$X_{non-Hispanic Whites}$ and $X_{Black}$ separately follow poisson distributions as followed: 

\begin{gather*}
X_{Asians}\thicksim Poi(0.464)\\
X_{non-Hispanic Whites}\thicksim Poi(21.901)\\
X_{Black}\thicksim Poi(0.640)
\end{gather*}

$$
X_{Asians}\thicksim Poi(0.464)\\
X_{non-Hispanic Whites}\thicksim Poi(21.901)\\
X_{Black}\thicksim Poi(0.640)
$$

Therefore
$$
P(X_{Asians}=30) = \frac{\lambda^xe^{- \lambda}}{x!} =  \frac{0.464^{30}e^{0.464}}{30!} = `r dpois(30,0.464)`\approx2.346 \times 10^{-43}
$$ 


$$
P(X_{non-Hispanic Whites}=30) = \frac{\lambda^xe^{- \lambda}}{x!} =  \frac{21.901^{30}e^{21.901}}{30!} = `r dpois(30,21.901)`\approx0.019
$$ 


$$
P(X_{Black}=30) = \frac{\lambda^xe^{- \lambda}}{x!} =  \frac{0.640^{30}e^{0.640}}{30!} = `r dpois(30,0.640)`\approx3.046 \times 10^{-39}
$$

As the result shows, a disparity (between the probability of exactly 30 uveal melanoma cases occurring in a given year) exists for different racial/ethnic groups. The non-Hispanic whites have the highest probability of 0.019, followed by Blacks and Asians, which have rather low probability.

# Problem2
## a)
### Intervention Group
The hypothesis is:
$$
H_0:\mu_1 - \mu_2=0  \quad vs \quad H_1:\mu_1-\mu_2   \neq 0 
$$
With the significance level $\alpha=0.05$, compute the test statistic: 
$$
t_{stats} = \frac{\bar{d}-0}{s_d/\sqrt{n}} = \frac{-0.76}{1.44/\sqrt{36}}=-3.17
$$
Because $|t_{stats}|=3.17 > t_{35,0.975} = `r round(qt(0.975,35),digits=2)`$, at 5% significance level, we reject $H_0$ and conclude that there's a difference of BMI between 6-months follow-up and baseline among intervention group.

### Control Group
The hypothesis is: 
$$
H_0:\mu_1 - \mu_2=0  \quad vs \quad H_1:\mu_1-\mu_2   \neq 0
$$
With the significance level $\alpha=0.05$, compute the test statistic:
$$
t_{stats} = \frac{\bar{d}-0}{s_d/\sqrt{n}} = \frac{0.28}{0.97/\sqrt{36}}=1.73 
$$
Because $|t_{stats}|=1.73 < t_{35,0.975} = `r round(qt(0.975,35),digits=2)`$, at 5% significance level, we fail to reject $H_0$ and conclude that there's no enough evidence to support a difference of BMI between 6-months follow-up and baseline among control group.

## b) 
Now perform a test to compare the BMI absolute changes between the two groups.

### Test for Equality of Variances
The hypothesis is:
$$
H_0:\sigma_1 = \sigma_2  \quad vs \quad H_1:\sigma_1 \neq \sigma_2  
$$
With the significance level $\alpha=0.05$, compute the test statistic:
$$
F_{stats} = \frac{s_1^2}{s_2^2} = \frac{1.44^2}{0.97^2}=2.204 \thicksim F_{35,35} 
$$
Because $F_{stats}=2.204 > F_{35,35,0.975}= `r round(qf(0.975,35,35),digits=3)`$ , at 5% significance level, we reject $H_0$ and conclude that there's a difference of variances of BMI change between intervention and control group. 

### Two-Sample Independent t-test with Unequal Variances
$$
H_0:\mu_1 = \mu_2  \quad vs \quad H_1:\mu_1 \neq \mu_2  
$$
With the significance level $\alpha=0.05$, compute the test statistic:
$$
t_{stats} = \frac{\bar{X_1}-\bar{X_2}}{\sqrt{\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2}}} = \frac{-0.76-0.28}
{\sqrt{\frac{1.44^2}{36}+\frac{0.97^2}{36}}}= 3.594
$$
$$
d'=\frac{(\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2})^2}{(\frac{s_1^2}{n_1})^2/(n_1-1)+(\frac{s_2^2}{n_1})^2/(n_2-1)} = \frac{(\frac{1.44^2}{36}+\frac{0.97^2}{36})^2}{(\frac{1.44^2}{36})^2/(36-1)+(\frac{0.97^2}{36})^2/(36-1)}= 61.340 = 61
$$
Because $t_{stats}= 3.594 > t_{61,0.975} = `r round(qt(0.975,61),digits=3)`$, at 5% significance level, we reject $H_0$ and conclude that there's a difference of BMI change between intervention and control group.

```{r, include = FALSE}
library(tidyverse)
df <- readxl::read_excel("./Exercise.xlsx",skip = 1) %>% 
        janitor::clean_names() %>% 
        rename(bmi_pre = pre_13,
               bmi_post = post_14
               ) %>% 
        select(bmi_pre,
               bmi_post,
               group
               ) %>% 
        mutate(group = factor(group,labels = c("Control","Intervention")),
               change = bmi_post - bmi_pre)

int_group <- filter(df,group == "Intervention")
ctr_group <- filter(df,group == "Control")

#P2_a conduct paired t-test 
t.test(int_group$bmi_post,int_group$bmi_pre, paired = T)
t.test(ctr_group$bmi_post,ctr_group$bmi_pre, paired = T)

#P2_b
## Test for variance
var.test(change~group, data = df)  # why the results are diff.!!!???
var.test(int_group$change,ctr_group$change)

##conduct Two-Sample Independent t-test with Equal Variances
t.test(change~group,data = df, var.equal = FALSE)  # the results are the same.
t.test(int_group$change,ctr_group$change,var.equal = FALSE)

```

## c) 
The assumption is that the distribution of BMI is normal.  

### i) Check the normality
```{r,echo=FALSE}
df %>% 
  pivot_longer(starts_with("bmi"),
               names_to = "time",
               values_to = "bmi") %>%
  mutate( time = recode(time,
                        bmi_pre = "baseline",
                        bmi_post = "6 months follow-up"),
          time = factor(time, levels = c("baseline","6 months follow-up"))
          ) %>% 
  ggplot(aes(x = bmi,fill = group)) +
  geom_density(alpha = .5) +
  facet_grid(time~.) +
  labs(title = "Distribution  of BMI at baseline and after 6 months" , x = "BMI" ) +
  theme(legend.position = "bottom", legend.title = element_blank(), plot.title = element_text(hjust = 0.5))
```
As we can see in the plot above, the distribution of BMI is right-skewed. In another word, the normality assumption is (very likely) invalid.  

### ii) Effect of non-normality and remedies
Our tests are based on the normality assumption, which decide the underlying distribution of BMI. Therefore, non-normality invalidate our tests and may distort the truth. 
Fortunately, we do have alternatives, such as **Non-parametric test** and **Transformation**.

# Problem3
Let X denote the number of restaurants that close by the end of 2019. As we know, 
$$
X\thicksim B(20,0.60)
$$
Therefore, use exact method:
$$
P(X \ge 10) = 1- P(X < 10) = 1- F(9) = `r round(1-pbinom(9,20,0.6),digits =3)`
$$

* Poisson approximation: since n=20 is not large and p=0.60 is not small, it's inappropriate to use poisson approximation to binomial. 
* Normal approximation: since np(1-p) = 4.8  $\approx$ 5, we may try to use normal approximation to binomial.

$$
X\thicksim  N(np,np(1-p)) =N(12,4.8)
$$

$$
P(X \ge 10) = 1- P(X < 10)  = `r round(1-pnorm(9,12,sqrt(4.8)),digits =3)`
$$

* As we can see, when npq is not greater than 5, it may not be very appropriate to use normal distribution to binomial. 
* According to the result, the probability that more than 10 restaurants will close by the end of 2019 is 87.2%.

# Problem4
## a)
### Paired t-test with Equal Variances
$$
H_0:\mu_{drug1} -\mu_{drug1}=0  \quad vs \quad H_1:\mu_{drug1} -  \mu_{drug2}  <0
$$
With the significance level $\alpha=0.05$, compute the test statistic:
$$
t_{stats} = \frac{\bar{d}-0}{s_d/\sqrt{n}} = \frac{-1.58}{1.23/\sqrt{10}}= -4.062 
$$
Because $t_{stats}=-4.06 < t_{9,0.05} = `r round(qt(0.05,9),digits=3)`$, at 5% significance level, we reject $H_0$ and conclude drug2 has a better efficacy than drug1 in increasing sleep time.

```{r,include = FALSE}
df4 <- 
        sleep %>%
        pivot_wider(
                names_from = group,
                values_from = extra
        ) %>% 
        mutate(change = `1` - `2`)
t.test(df4$`1`,df4$`2`,alternative = "less",paired = TRUE)
t.test(df4$change,alternative = "less",mu = 0)

```

## b)
According to the formula
$$
95 \% CI = (-\infty,\bar{d}+t_{1-0.05,9}S/\sqrt{n})
$$
$$
95 \% CI = (-\infty,-1.58+1.833\times 1.230/\sqrt{10}) =(-\infty,-0.867)
$$

## c)
According to the formula
$$
Power = 1 -\beta = P(reject H_0|H_1) = F_t(t_\alpha+ \frac{|\Delta|}{\sigma/ \sqrt{n}}) = 
F_t(-1.83+ \frac{1.58}{1.230/\sqrt{10}}) = 0.974
$$

* The posterior power is 0.974, meaning that if there's anything significant, the probability that we can detect it is 97.4%.

```{r,include = FALSE}
power.t.test(10 ,delta = 1.58/1.230, sd = 1.230, alternative = "one.sided",type = "one.sample") ### diff???
```

## d) PROs and CONs of using a posteriori/post-hoc power analysis 

* PROs: 
When already having a fixed sample size, we can use posteriori power analysis before conducting the study to see
whether or not such study would give us adequate power.

* CONs:
Post-hoc power analysis may been misused in interpreting nonsignificant study results, where there will be always a low power. As a result, an observed nonsignificant result may be misinterpreted as the study having inadequate power(underpowered to detect) but as a fact, the result may come from the true nonsignificance.
