---
title: "p8130_hw3_xj2249"
author: "xj2249"
date: "2019/10/24"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(pander)
library(multcomp)
```

# Problem2
## a)
```{r}

sur_df <- read_csv("./hw3/SurvCancer.csv") %>% 
        janitor::clean_names() %>% 
        mutate(organ = str_to_lower(organ),
               organ = factor(organ))
# boxplot
sur_df %>% 
        ggplot(aes( x =  organ, y = survival)) +
        geom_boxplot()

# descriptive statistics 
sur_df %>% 
        group_by(organ) %>% 
        summarise( n = n(),
                   mean = mean(survival),
                   median = median(survival),
                   sd = sd(survival),
                   IQR = IQR(survival),
                   min = min(survival),
                   max = max(survival)
                  ) %>% 
        knitr::kable()
```

## b)
Hypothesis:
$$
H_0:\mu_1=\mu_2=\mu_3=\mu_4=\mu_5 \quad vs \quad H_1: 
\text{not all means are equal}
$$
Significance level:$\alpha = 0.01$   

Assumptions:(1)Independence (2)equal variance (3)normality   

Decision rule:
Reject $H_0$: if $F_{stats} > F_{4,59,1-\alpha/2}= F_{4,59,0.995}=$ `r round(qf(0.995,4,59),digits =3)`
Fail to reject $H_0$: if $F_{stats} < F_{4,59,0.995}$

Interpretation:
Since $F_{stats} = 6.433  > F_{4,59,1-\alpha/2}= F_{4,59,0.995}=$ `r round(qf(0.995,4,59),digits =3)`, we reject $H_0$ and conclude that there is a significant difference in average survival time among different cancer groups.

```{r}
sur_aov <- aov(survival ~ organ, data = sur_df)

pander(sur_aov)

```

## c) how to adjust 0.05 to 0.01
1) Bonferroni
```{r}
pairwise.t.test(sur_df$survival, sur_df$organ, p.adj = 'bonferroni')  
```

2) Tukey
```{r}
TukeyHSD(sur_aov,conf.level = 0.99) ## 0.99???
TukeyHSD(sur_aov,conf.level = 0.99) %>% plot()

glht(sur_aov, linfct = mcp(organ = "Tukey")) %>% summary()
glht(sur_aov, linfct = mcp(organ = "Tukey")) %>% plot()
```

3) Dunnett Test
```{r}
glht(sur_aov, linfct = mcp(organ = "Dunnett")) %>% summary()
```

## d) 
i) check the normality assumption
```{r}
# first, try a hist/density plot.
hist(residuals(sur_aov)) 

# check the normality (of residuals) assumption
qqnorm(residuals(sur_aov))

```
As the QQ-plot shows, the normality assumption is questionable. Therefore, we can  Kruskal Wallis test to fix the problem.  
ii) KW test 
```{r}
kruskal.test(survival ~ organ, data = sur_df) %>% pander()
```
The p-value is 0.005, at a significance level 0.01, we reject $H_0$ and conclude that there is a significant difference in average survival time among different cancer groups. The p-value of kw-test is 0.004798, much larger than that of the anova test(0.0002295), which shows that kw-test is harder to reject $H_0$ and it's more conservative and less powerful. 

# Problem3
