---
title: "p8130_hw3_xj2249"
author: "xj2249"
date: "2019/10/24"
output: pdf_document
---
  
```{r setup, include=FALSE}
library(tidyverse)
library(multcomp)
library(arsenal)
library(kableExtra)

theme_set(theme_minimal() +  theme(plot.title = element_text(hjust = 0.5)))


setwd("/Users/xuejin/Desktop/CUMC_semester1/Courses/Biostatistics Methods I/P8130_hw_xj2249")
sur_df <- read_csv("./hw3/SurvCancer.csv") %>% 
        janitor::clean_names() %>% 
        mutate(organ = str_to_lower(organ),
               organ = factor(organ))
```

# Problem2
## a)
```{r descriptive statistics and boxplot}
# descriptive statistics 
control_table1 <- tableby.control(
        total = FALSE,
        test = FALSE,
        numeric.stats = c("meansd","medianq1q3","range"),
        stats.labels = list(meansd = "Mean (SD)",
                            medianq1q3 = "Median (Q1, Q3)",
                            range = "Min - Max"),
        digits = 2
        ) 
tab1 <- tableby(organ~survival,sur_df,control = control_table1)

summary(tab1,text = TRUE) %>% kable(caption = "Descriptive statistics for each group") %>% kable_styling(latex_options = "scale_down")

sur_df %>% 
        ggplot(aes( x =  organ, y = survival , fill = organ)) +
        geom_boxplot() + 
        labs( y = "survival times (days)",
              title = "survival times by group") 
```

As we can see, average survival time varies among different cancers. Breast cancer group has the largest mean and standard deviation value, followed by ovary, colon, stomach and bronchus cancer groups. 

## b)
```{r, echo = F}
# Anova table
sur_aov <- aov(survival ~ organ, data = sur_df)
pander::pander(sur_aov)
```

1) Hypothesis:
$$
H_0:\mu_1=\mu_2=\mu_3=\mu_4=\mu_5 \quad vs \quad H_1: 
\text{not all means are equal}
$$
2) Significance level: $\alpha = 0.01$   

3) Assumptions:(i)Independence (ii)equal variances (iii)normality   

4) Decision rule:  
Reject $H_0$: if $F_{stats} > F_{4,59,1-\alpha}= F_{4,59,0.99}=$ `r round(qf(0.99,4,59),digits =3)`  
Fail to reject $H_0$: if $F_{stats} < F_{4,59,0.99}$

5) Interpretation:
Since $F_{stats} = 6.433  > F_{4,59,1-\alpha}= F_{4,59,0.99}=$ `r round(qf(0.99,4,59),digits =3)`, we reject $H_0$ and conclude that there is a significant difference in average survival time among different cancer groups.


## c) pairwise comparisons
1) Bonferroni
```{r}
pairwise.t.test(sur_df$survival, sur_df$organ, p.adj = 'bonferroni', conf.level = 0.99) 
```

2) Tukey
```{r fig.width=8, fig.height=4}
TukeyHSD(sur_aov,conf.level = 0.99) 
par(mar=c(2,8,2,2))
TukeyHSD(sur_aov,conf.level = 0.99) %>% plot(las = 1)
```

3) Dunnett Test
```{r}
glht(sur_aov, linfct = mcp(organ = "Dunnett"),conf.level =0.99) %>% summary()
```

The main conclusions of these three methods are consistent with each other: at a significance level of 0.01, we can conclude that there's a significant difference in average survival time between bronchus and breast cancer; colon and breast cancer; stomach and breast cancer. According to Bonferroni and Tukey, we can also conclude that there's no there's no enough evidence to support a significant difference between other pairwise groups.  
As we can see, the p-value using Tukey's method is smaller than that of Bonferroni's methods, indicating than Tukey is less conservative than Bonferroni. And Dunnett is used to compare with a specific group rather than any pairwise comparisons, like Bonferroni and Tukey. 

## d)  
i) check the normality assumption
```{r}
# first, try a hist/density plot.
hist(residuals(sur_aov)) 

# check the normality (of residuals) assumption
qqnorm(residuals(sur_aov))
qqline(residuals(sur_aov),lwd = 2)
```
 
As the QQ-plot shows, the normality assumption is questionable. Therefore, we can  use **non-parametric test (Kruskal Wallis test)** or **transformation** to fix the problem. 

ii) KW test 
```{r}
kruskal.test(survival ~ organ, data = sur_df) %>% pander::pander()
```
  
Since p-value < 0.01, at a significance level 0.01, we reject $H_0$ and conclude that there is a significant difference in average survival time among different cancer groups. The p-value of kw-test is 0.004798, much larger than that of the anova test(0.0002295), which shows that kw-test is harder to reject $H_0$ and it's more conservative and less powerful. 

# Problem3
a) 
$$
\begin{aligned}
Average_{zinc} &=\frac{n_{zinc}\times Avg_{zinc}+n_{zinc+iron}\times Avg_{zinc+iron}}
{n_{iron}+n_{zinc+iron}}\\
&= \frac{54 \times 0.7+55\times0.8}{54+55}\\
&= `r  round((54*0.7+55*0.8)/(54+55),digits=2)`
\end{aligned}
$$
$$
\begin{aligned}
Average_{non-zinc} &=\frac{n_{placebo}\times Avg_{placebo}+
n_{iron}\times Avg_{iron}}
{n_{placebo}+n_{iron}}\\
&= \frac{56 \times 1.1+54\times1.4}{56+54}\\
&= `r  round((56*1.1+54*1.4)/(56+54),digits=2)`
\end{aligned}
$$
$$
diff = Average_{zinc}-Average_{non-zinc}=`r  round((54*0.7+55*0.8)/(54+55)-(56*1.1+54*1.4)/(56+54),digits=2)`
$$
b)
$$
\begin{aligned}
&sd_{placebo} =s.e._{placebo}\times \sqrt{n_{placebo}}=\sqrt{56} \times 0.2 =`r round(sqrt(56)*0.2,digits = 2)`\\
&sd_{iron} =s.e._{iron}\times \sqrt{n_{iron}}=\sqrt{54} \times0.2 =`r round(sqrt(54)*0.2, digits = 2)`\\
&sd_{zinc} =s.e._{zinc}\times \sqrt{n_{zinc}}=\sqrt{54} \times 0.1 =`r round(sqrt(54)*0.1, digits = 2)`\\
&sd_{zinc+iron} =s.e._{zinc+iron}\times \sqrt{n_{zinc+iron}}=\sqrt{55} \times 0.1 =`r round(sqrt(55)*0.1,digits = 2)`\\
\end{aligned}
$$

Between 4 initial groups, placebo and iron group share similar and larger standard deviation, zinc and zinc+iron group share similar and smaller standard deviation. 

To decide if we can assume equal variances, we can use F-test. For "zinc" and "zinc-iron", $F_{stats} =\frac{s_{placebo}^2}{s_{zinc+iron}^2} =\frac{0.73^2}{0.74^2}=0.97 \leq F_{53,54,1-0.05/2}$ = `r round(qf(0.975,53,54),digits = 2)` and $\geq F_{53,54,0.05/2}$ = `r round(qf(1-0.975,53,54),digits = 2)`. For "placebo" and "iron" groups, $F_{stats} =\frac{s_{zinc}^2}{s_{iron}^2} =\frac{1.50^2}{1.47^2}=1.04 \leq F_{55,53,1-0.05/2}$ = `r round(qf(0.975,55,53),digits = 2)` and  $\geq F_{55,53,0.05/2}$ = `r round(qf(1-0.975,55,53),digits = 2)`  

Therefore, we can assume that the standard deviation is equal b/w "zinc" and "zinc-iron", and also b/w "placebo" and "iron".

$$
\begin{aligned}
s^2_{zinc} &= \frac{s^2_{zinc}\times(n_{zinc}-1)+s^2_{zinc+iron}\times(n_{zinc+iron}-1)}
{n_{iron}+n_{zinc+iron}-2}
\\
&=\frac{0.73^2 \times53 +0.74^2\times 54}{54+55-2}
\\
&= 0.54
\\
s^2_{non-iron} &=\frac{s^2_{placebo}\times(n_{placebo}-1)+s^2_{iron}\times(n_{iron}-1)}
{n_{placebo}+n_{iron}-2}\\
&=\frac{1.50^2 \times55 +1.47^2\times 53}{56+54-2}
\\
&= 2.21
\end{aligned}
$$

c)
i) Equal allocation
$$
\begin{aligned}
n &=\frac{\left(\sigma_{zinc}^{2}+\sigma_{non-zinc}^{2}\right)\left(z_{1-\alpha / 2}+z_{1-\beta}\right)^{2}}{\Delta^{2}}\\
&=\frac{(0.54+2.21)(1.96+0.84)^2}{0.5^2}\\
&=86.24\\
&\approx87
\end{aligned}
$$

Therefore, the sample size is 87 for both zinc and non-zinc group.

ii) 2:1 allocation
$$
k = n_{non-zinc}/n_{zinc}  = 2
$$

$$
\begin{aligned}
n_{zinc}&=\frac{\left(\sigma_{zinc}^{2}+\sigma_{non-zinc}^{2} / k\right)\left(z_{1-\alpha / 2}+z_{1-\beta}\right)^{2}}{\Delta^{2}} \\
&=\frac{(0.54+2.21/2)(1.96+0.84)^2}{0.5^2}\\
&=51.59\\
&\approx52
\end{aligned}
$$
Therefore, the sample size for zinc group is 52. 
$$
\begin{aligned}
n_{non-zinc}&=\frac{\left(k \sigma_{zinc}^{2}+\sigma_{non-zinc}^{2}\right)\left(z_{1-\alpha / 2}+z_{1-\beta}\right)^{2}}{\Delta^{2}}\\
&= \frac{(2\times0.54+2.21)(1.96+0.84)^2}{0.5^2}\\
&= 103.17\\
&\approx 104
\end{aligned}
$$
Therefore, the sample size for non-zinc group is 104. 



