---
title: "rec11"
author: "xj2249"
date: "2019/11/19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Problem1
```{r}
brand = tibble(
    y = c(64,73,61,76,72,80,71,83,83,89,86,93,88,95,94,100),
    x1 = c(4,4,4,4,6,6,6,6,8,8,8,8,10,10,10,10),
    x2 = c(rep(c(2,4),8))
)

model <- lm(y ~.,data = brand)
summary(model)
```

## a)
studentized residual(internal)
```{r}
stu_res<-rstandard(model)
outliers_y <- stu_res[abs(stu_res)>2.5]
```
no outlier in y

## b)
Leverage values
```{r}
influence.measures(model) %>% as_tibble()

```
no

## c)
```{r}
hatvalues(model)
sort(stu_res)
case14 <- influence.measures(model) %>% tibble() %>% pull(.) %>% .[[1]] %>% .[14,]
case14 
```

cut off 0.2  moderate leverage catogory
or 2p/n

abs(dffit) = 1.17 , 
cook < 0.5 
do not know!!! possible?

potential point so need to check it out 

## d)
```{r}
# with case 14
model <- lm(y ~.,data = brand)
# without case 14
model_omit <- lm(y ~.,data = filter(brand, y !=95))
# brand2 = brand[-14,]

summary(model)
summary(model_omit)
```

coeficient increaase, and se decrease !
changed, but only moderatly(<10%)

# Problem2
```{r}
hos <- 
    read_csv("Hospital.csv") %>% 
    janitor::clean_names() %>%
    dplyr::select(-id,-xray,-beds,-medschl,-region)

names(hos) <- c("y","x1","x2","x3","x4","x5","x6")
```

## new tool from rec10_solution code!!! looks good!
```{r}
DataExplorer::plot_histogram(hos)
DataExplorer::introduce(hos)
```

##new color "coral"
```{r}
hos %>% 
    ggplot(aes(x = x1, y = y )) +
    geom_smooth(color = "coral")
```


## a) Scatter plot matrix for all variables
```{r}

hos %>% 
    pairs()

cor <- 
    hos %>% 
    cor() %>% 
    round(digits = 2) 

```

x4 and x5, x5 and x6 , x4 and x6 are corrlated


## b)
```{r}
anova(lm(y~x1 + x2, data = hos))
```
SSE(x1,x2) = 278.250 

```{r}
anova(lm(y~x1 + x2 +x3, data = hos))
anova(lm(y~x1 + x2 +x4, data = hos))
anova(lm(y~x1 + x2 +x5, data = hos))
anova(lm(y~x1 + x2 +x6, data = hos)) %>% broom::tidy()
```

SSR(x3|x1,x2) = 3.248
SSR(x4|x1,x2) = 37.899 
SSR(x5|x1,x2) = 10.397
SSR(x6|x1,x2) = 10.125 

R^2(x3|x1,x2) = 3.248/278.250 = 0.01167296
R^2(x4|x1,x2) = 37.899/278.250 = 0.1362049
R^2(x5|x1,x2) = 10.397/278.250 =  0.03736568
R^2(x6|x1,x2) = 10.125/278.250 =  0.03638814

```{r}
3.248/278.250 
37.899/278.250 
10.397/278.250
10.125/278.250 
```

## c)
x4 is the best  
SSR(x4|x1,x2) = 37.899  yes, largest

## d)
```{r}
model12 = lm(y~x1 + x2,data = hos)
model124 = lm(y~x1 + x2 +x4, data = hos)
anova(model12,model124)
(37.899/1)/(2.205)
summary(model12)
anova(model12)
summary(model124)
library(HH)
vif(model124)
summary(lm(x1 ~ x2 +x4, data = hos))
1/(1-0.003566) #that's how vif is caculated!!!
```
all less than 5, so no multi...

## e)
# automatic 
```{r}
library(ledp)
model_full <- lm(y ~ ., data = hos)
summary(model_full)
backward1 <- update(model_full,.~.,-x6)
summary(backward1)

step(model_full, direction='backward')
# leaps for mallow cp $adjr2 give adj r-squared 
```



# let's try to verify something
##  SSR(X1) in only x1 and SSR(X1) in x1 and x2  
```{r}
# SSR(X1) in only x1 and SSR(X1) in x1 and x2  
anova(lm(y~x1, data = hos)) 
#SSR(X1) = 14.60 in x1
anova(lm(y~x1 + x2 , data = hos))
# SSR(X1) in x1 and x2  = 14.604 
# OMG same? 

# try other
anova(lm(y~x1 + x2 + x3, data = hos))

# yes! SSR(X2|X1) is the same too

# sequence matters!
anova(lm(y~x3 + x2 + x1, data = hos))
summary(lm(y~x3 + x2 + x1, data = hos))
summary(lm(y~ x1 +x3 + x2 ,data = hos))
```

## partial R^2
```{r}
cor  
```
x4 and x5 are highly correlated, let's see what's gonna happen 

# change in coefficient and se
```{r}
summary(lm(y~x4, data = hos))
# beta= 0.005891  and  se= 0.001039
# R-squared:  0.2246,	Adjusted R-squared:  0.2176 
summary(lm(y~x4+x5, data = hos))
# beta= 0.011663  and  se= 0.002415
# R-squared: 0.2705,	Adjusted R-squared:  0.2573
summary(lm(y~x5, data = hos))
```
coeficient change, se increase and Adjusted R-squared decrease, R-squared increase.

# paritial R-square of x4  and cody's way to calculate vif
```{r}
anova(lm(y~x5, data = hos))
anova(lm(y~x5+x4, data = hos))
63.298/361.80
1/(1- 0.174953)
```
SSE(X5) = 361.80
SSR(X4|X5) = 63.298
R^2(X4|X5) = 63.298/361.80 = 0.174953 
vif(?) = 1.212052 that's not right. cody's way is not working.

annother
```{r}
summary(lm(x4~x5, data = hos))
vif(lm( y ~x4+x5, data = hos))
```

vif = 1/(1-0.8243) = 5.69152

## partial t-test, namely all t-tests in summary are testing marginal significance of xi afrer other predictors have been taken into account
```{r}
summary(lm(y~x1+x2+x3, data = hos))
anova(lm(y~x1+x2+x3, data = hos))
anova(lm(y~x1+x3 +x2, data = hos))

# the t value will be diff from the t-value with x2+x3 in. 
summary(lm(y~x1, data = hos))
1.135^2
4.908^2
```
for x3, 1.135^2 = 1.28 = F
for x2, 4.908^2 = 24.09 = F




